---
title: "Project 1 - Loan"
author: Hye Young Park
date: June 16th, 2019
output: html_document
---
Source: https://www.kaggle.com/wendykan/lending-club-loan-data

## Load the Library
```{r}
library(dplyr)
```

Here is data regarding loans issued from "lending-club-loan-data". First of all, let's look at all the columns of data.

## Load the Sheet
```{r}
loan<-read.csv("C:/Users/illya/Downloads/loan.csv")
```

## Dimensions and Structure of Dataset
```{r}
dim(loan)
str(loan)
```

## First 6 Complete Rows and Last 6 rows
```{r}
head(loan)
tail(loan)
```

## Removing Missing Values

To make the data tidy, we are going to remove NA values. There are two ways to do that. One of them is using loan[complete.cases(loan),] and the other one is using na.omit(loan).
Note: na.omit removes the rows with NA in them but preserves the row numbers

```{r}
loan_complete <- loan[complete.cases(loan),]
dim(loan_complete)
loan_complete <- na.omit(loan)
dim(loan_complete)
tail(loan_complete)
```

## Organizing the Data

We are looking at the first few columns in order to get a better feel for the data. Besides, the column of funded_amnt will be removed because the value for funded_amnt is same as the value of loan_amnt. We're also removing other columns
as they're irrelevant.

```{r}
loan_complete1 <- select(loan_complete, loan_amnt:annual_inc)
loan_complete1$funded_amnt <- NULL
loan_complete1$installment <- NULL
loan_complete1$sub_grade <- NULL
loan_complete1$grade <- NULL
head(loan_complete1)
```

## Removing Small Loans (< 10K)

There is a lot of data, so we're trimming it down to
loan amounts greather than 10 thousand dollars to take a look
at some things closer.

```{r}
filtered_loan_amnt <- filter(loan_complete1, loan_amnt > 10000)
head(filtered_loan_amnt, 30)
dim(filtered_loan_amnt)
```

## Subsetting

Let's take a look at loans for which the term is 36 months

```{r}
term <- subset(filtered_loan_amnt, term == " 36 months")
str(term)
```

Take a look at annual income below 60k 

```{r}
below_sixty <- subset(term, annual_inc < 60000)
dim(below_sixty)
```

and above 60k to see
the different populations who took out 36 month loans

```{r}
above_sixty <- subset(term, annual_inc > 60000)
dim(above_sixty)
```


## Reordering the Data

We can arrange the data to be sorted by loan amount and look at the
bottom 10 values

```{r}
head(loan_complete1, 10)
loan_amnt_organized <- arrange(filtered_loan_amnt, desc(loan_amnt))
tail(loan_amnt_organized, 10)
```