---
title: "Project 3 - Loan"
author: Hye Young Park
date: June 30th, 2019
output: html_document
---
  Source: https://www.kaggle.com/wendykan/lending-club-loan-data

Here is data regarding loans issued from "lending-club-loan-data". 
We are interested in seeing if the loan amount is affected
by annual income, and if so, how?
  
```{r}
library(dplyr)
library(package = "lattice")
library(ggplot2)
```

## 1.	Formulate your question

Following the EDA checklist, lets formulate our question.
Does annual income affect the size of the loan amount funded?

## 2.	Read in your data

```{r}
loan<-read.csv("C:/Users/illya/Downloads/loan.csv")
```

## 3.	Check the packaging
```{r}
nrow(loan)
ncol(loan)
```


## 4.	Run str()
```{r}
str(loan)
```

## 5.	Look at the top and the bottom of your data
```{r}
head(loan)
tail(loan)
```

## 6.	Check your n’s (the count)
```{r}
#what loan amounts
table(loan$loan_amnt)

#how many different incomes?
select(loan, annual_inc) %>% unique %>% nrow 
```


## 7.	Validate with at least one external data source
```{r}
#check standards
summary(loan$annual_inc)

#look at deciles of the data
quantile(loan$annual_inc, seq(0, 1, 0.1)) 
```

## 8.	Try the easy solution first
```{r}
ranking <- group_by(loan, loan_amnt) %>%  
  summarize(Median.Income = median(annual_inc)) %>% 
  as.data.frame %>%  
  arrange(desc(loan_amnt))
head(ranking,10)

#one outlier (in terms of income) who took out a large loan 
filter(loan, loan_amnt == 39200) %>% nrow

#number of people who took out the highest loan
filter(loan, loan_amnt == 40000) %>% nrow

tail(ranking,10)
```

## 9.	Challenge your solution
Not necessary as easy solution was good enough for our purposes.

## 10. Follow up
We found that there was a tendency for people with larger incomes to take out larger
loans in terms of amount.

## Ggplot 2 graphs and visualization

Next well look at some data regarding loan term and interest rates.
This will show that people tend to take out smaller loans for shorter terms.
We see that 50% of the people who took a 36 month loan had the Loan Amount < 10,000.
We also see that no one who took out a 60 month loan took a loan that was < 10,000.

```{r}
boxplot(loan_amnt ~ term, loan, xlab="Loan Term", ylab="Loan Amount",col.axis="blue",col.lab="red")
```


Here we see a histogram of the annual income. Its skewed right so we use median as our
measure of average income in the previous sections.

```{r}
hist(loan$annual_inc, breaks = 100, col = "blue")
```

Here is a fun plot showing the amount of loans by term.

```{r}
qplot(loan_amnt, data = loan, binwidth=1000, fill = term)
```

Here are two graphs, one for each term. They show the histograms for the interest rate for each of the terms
```{r}
qplot(int_rate, data = loan, facets = .~term, fill = term) + ggtitle("Loan Interest by term")
```

A simple bar chart of how many loans were taken out in each term.

```{r}
ggplot(data = loan) + 
  geom_bar(mapping = aes(x = term, fill= term))
```


------------------ PROJECT 3 ------------------

Null hypothesis - using categorical value "term" and quantitative variable "amount"
I suspect it will be rejected due to variability between interest rates and incomes
for the different loan terms.
These are our data, loans with 36 month rates and loans with 60 month rates.

```{r}
qplot(loan_amnt, data = loan, binwidth=1000, fill = term)
```

We want to get rid of outliers in the 36 month dataset, since
they heavily skew the graph to the right. First we will filter the dataset
to only have 36 month loans, then filter it to have loans that are in the IQR range, getting
rid of outliers, then take a random sample so that it matches the number of loans that
are 60 months in their term.
```{r}
loan36 <- filter(loan, loan$term == " 36 months") ## get 36 month loans
thirtySixFiltered<-loan36[loan36$loan_amnt < 20000,] ##normalize data, get rid of outliers
loanLesstwenty <- sample_n(thirtySixFiltered, 3100)##match "n" of loan60
```
summary
```{r}
summary(loanLesstwenty$loan_amnt)
```
Standard Deviation
```{r}
sd(loanLesstwenty$loan_amnt)
```
a boxplot :)
```{r}
boxplot(loanLesstwenty$loan_amnt, main = "Loan Amount less than 20k", col = "blue")
```

We do the same for the 60 month term loans.
```{r}
loan60 <- filter(loan, loan$term == " 60 months") 
```
We don't filter it since the data looks normal without many outliers.

Summary
```{r}
summary(loan60$loan_amnt)
```
standard deviation
```{r}
sd(loan60$loan_amnt)
```
boxplot
```{r}
boxplot(loan60$loan_amnt, main = "Loan Amount - Term: 60 Month", col = "blue")

loanSixtyAmnt <- loan60$loan_amnt
loanThirtySixAmnt <- loanLesstwenty$loan_amnt
```
Next we run the t test on the resulting vectors, which will show us
that we reject the null hypothesis, there IS a large difference in the means
and I suspect that interest rate and annual income play a role in that.
```{r}
t.test(loanSixtyAmnt, loanThirtySixAmnt)
```


